---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

### - Show any code that is needed to Load the data (i.e. read.csv())

```{r,echo=TRUE,cache=TRUE}
a=read.csv("~/activity.csv")

```


## What is mean total number of steps taken per day?
### 1. Calculate the total number of steps taken per day
```{r,echo=TRUE}
b=aggregate(a$steps,by=list(a$date),sum)
names(b)=c("date","steps") 
b
```



## What is the average daily activity pattern?
### 2. Make a histogram of the total number of steps taken each day
```{r,echo=TRUE}
hist(b$steps,col="red",xlab="steps",main="histogram of total  steps taken per day with missing value")
```

### 3. Calculate and report the mean and median of the total number of steps taken per day

```{r,echo=TRUE,results='hide'}
day_mean=mean(b$steps,na.rm=T)
## force to express result without scientific notation with option scipen
day_mean=format(day_mean,scipen=100,digits=5)
day_median=median(b$steps,na.rm=T)
```
### The mean of days is `r day_mean`

### The median of days is `r day_median`

## What is the average daily activity pattern?

### Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r,echo=TRUE}
library(reshape2)
md=melt(a,id.vars=c("date","interval"),na.rm=T)
## create new dataframe with 2 column -interval and mean steps
new=dcast(md,interval~variable,mean)
## make the plot
plot(new,type="l",col="red",main=" average steps of 5-minute interval")
```



## Imputing missing values

```{r}
## dataframe nof subset of record contain missing value
nof=a[is.na(a$steps)==T,]
number=nrow(nof) ## total records that contain NA
```

### 1.The total number of missing values in the dataset is `r number` 

### 2. Devise a strategy for filling in all of the missing values in the dataset using the mean/median for that day, or the mean for that 5-minute interval, etc.

### **Missing value, will filling with  mean of that 5-minute interval**

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r,echo=TRUE,results='hide'}
##let dataframe a add a 5-minute interval average steps column
library(dplyr)
a=tbl_df(a)
new=tbl_df(new)
extstep=new 
for(i in 1:60){extstep=rbind(extstep,new)}
names(extstep)=c("interval","mean")
exta=cbind(a,extstep$mean)## create a dataframe exta=original original dataframe + mean interval column
names(exta)=c("steps","date","interval","mstep")
## extab- subset of missing value records
extab=filter(exta,is.na(steps))
##replace missing steps with mean
extab=mutate(extab,steps=mstep)
extab=select(extab,steps,date,interval)
## isa - subset of complete records
isa=filter(a,!is.na(steps))
##rbind two subset to complete the whole data
data=arrange(rbind(isa,extab),date,interval)
##show the results of replacement
head(data)
```
### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r,echo=TRUE,results='hide'}
##grouping data by date
data_mean=group_by(data,date)
##sum up steps by date
sumsteps=summarise(data_mean,sum(steps))
names(sumsteps)=c("date","total")
##plot histgram
hist(sumsteps$total,main = "histogram of Total steps taken per day filling missing value with interval mean",xlab = "steps")
##calculate with mean and median 
filled_mean=mean(sumsteps$total)
## formatting the mean and median not to show as scientific format

filled_mean=format(filled_mean,scipen=100,digit=5)
filled_median=median(sumsteps$total)
filled_median=format(filled_median,scipen=100,digits = 5)
```
### the mean  total number of steps taken per day is `r filled_mean`

### the median total number of steps taken per day is `r filled_median`


## Are there differences in activity patterns between weekdays and weekends?
### For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

### Create a new factor variable in the dataset with two levels - "weekday and weekend"  indicating whether a given date is a weekday or weekend day.
```{r,echo=TRUE}
##change environment from Chinese to English
Sys.setlocale(,"USA")

data=mutate(data,week= weekdays(as.Date(date)) %in% c("Saturday" , "Sunday"))  
data$week=factor(data$week)
levels(data$week)=c("weekday","weekend")
Sys.setlocale(,"CHT")## change back to Chinese environment with "CHT"
```

### Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r,echo=TRUE}
fff=aggregate(data$steps,by=list(data$interval,data$week),mean)
names(fff)=c("interval","wday","steps")
library(ggplot2)
g=ggplot(fff,aes(interval,steps))
g+geom_line()+facet_grid(wday~.)+labs(title="Average steps of Weekday vs. Weekend")
